Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
Responded to some comments in Section 7

% begin related-work
\section{Relationship To Classic Symmetric Lenses}
\label{sec:relationship}
\bcp{I suggest adding an adjective---e.g., ``classic symmetric lenses''---to
avoid confusion between the old ones and our new ones.}
\paragraph*{Formalization} We have reformulated the problem addressed by
symmetric lenses with simple symmetric lenses. Classically, these types of
synchronization tasks are represented as symmetric lenses. A symmetric lens
$\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
designated element $init \in C$, and two functions for propagating changes in
each format to the other:
\begin{enumerate}
\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
\end{enumerate}
In this formulation, the components of each format that aren't shared with the
other format are stored in the complement. When one format is edited, the \PutR
or \PutL function stitches together that edited data with the data for the other
side, stored in the complement.  The designated element, $init$, is used as the
initial value of $C$, and specifies default behavior when there isn't existing
data to stitch with.  In our example\bcp{which example?}, the complement
would consist of two lists 
of commands, where one list contains the Linux commands and one list contains
the Windows commands.

@@

\paragraph*{Issues with Symmetric Lenses} Using symmetric lenses for
synchronization forces us to maintain not just two files that we aim to
sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
  should be careful not to oversell the difficulty of storing the
  complements.  It requires more engineering, but it's not rocket
  science.}This is unintuitive, and creates many problems in 
practice. For example, consider the situation where I am synchronizing a crontab
and job manager file. I am synchronizing across computers, so I must either
maintain a copy of the complement on each computer, and keep them synchronized,
or figure out a computer to hold the complement, and make sure all edits go
through that complement. Furthermore, if I want to change the synchronization to
a different computer, I don't need to merely make sure that my files are moved
to the computer, I also have to make sure that I correctly move the complement,
and hook up that complement to be used for that synchronization task. This is a
huge divergence from other synchronization utilities\bcp{No.  For example,
  Unison maintains a database on each replica recording, for each file, what
its hash was after the last time the two replicas were synchronized.}, where all
that needs to be 
done is maintain the files. Using those for synchronization, edits can be sent
over the wire, and the appropriate function can stitch together those edits, no
additional state needed.

For synthesis, because each lens has a custom complement, we can no longer
specify the put functions though mere input/output examples.  Instead, we have
to specify the behavior through the observable behavior of edit sequences.
Furthermore, there are lenses that require arbitrarily long edit sequences to
differentiate between: consider the lens that keeps two formats equal for the
first $k$ non-identity edits, then keeps them completely unsynchronized for all
future edits.  This requires an edit sequence of $k$ edits to differentiate
between this lens and the identity lens.  Simple symmetric lenses avoid this
problem by not allowing such lenses: we can't keep track of the number of
non-identity edits performed unless there are explicit fields in the data
formats keeping track of them.  While we are reducing the expressivity of our
lenses with this restriction, we feel the benefits outweigh the costs.

@@

Despite being less expressive than symmetric lenses, simple symmetric lenses are
strictly more expressive than classical asymmetric lenses.

\begin{theorem}\bcp{To prove this, we need to state the type and laws for
  asymmetric lenses.}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem}

>>>>

% begin related-work
\section{Relationship To Classic Symmetric Lenses}
\label{sec:relationship}
\paragraph*{Formalization} We have reformulated the problem addressed by
symmetric lenses with simple symmetric lenses. Classically, these types of
synchronization tasks are represented as symmetric lenses. A symmetric lens
$\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
designated element $init \in C$, and two functions for propagating changes in
each format to the other:
\begin{enumerate}
\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
\end{enumerate}
In this formulation, the components of each format that aren't shared with the
other format are stored in the complement. When one format is edited, the \PutR
or \PutL function stitches together that edited data with the data for the other
side, stored in the complement.  The designated element, $init$, is used as the
initial value of $C$, and specifies default behavior when there isn't existing
data to stitch with.  In the example in Figure~\ref{fig:minimized-representations} the complement would consist of a list of pairs with each pair consisting of an integral salaray value and the name of a company.

@@

\paragraph*{Issues with Symmetric Lenses}
\iffalse
Using symmetric lenses for
synchronization forces us to maintain not just two files that we aim to
sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
  should be careful not to oversell the difficulty of storing the
  complements.  It requires more engineering, but it's not rocket
  science.}This is unintuitive, and creates many problems in 
practice. For example, consider the situation where I am synchronizing a crontab
and job manager file. I am synchronizing across computers, so I must either
maintain a copy of the complement on each computer, and keep them synchronized,
or figure out a computer to hold the complement, and make sure all edits go
through that complement. Furthermore, if I want to change the synchronization to
a different computer, I don't need to merely make sure that my files are moved
to the computer, I also have to make sure that I correctly move the complement,
and hook up that complement to be used for that synchronization task. This is a
huge divergence from other synchronization utilities\bcp{No.  For example,
  Unison maintains a database on each replica recording, for each file, what
its hash was after the last time the two replicas were synchronized.}, where all
that needs to be 
done is maintain the files. Using those for synchronization, edits can be sent
over the wire, and the appropriate function can stitch together those edits, no
additional state needed.
\fi
For one, using classical symmetric lenses for synchronization forces us to maintain three files instead of two, which requires more engineering.
The more significant issues with classical symmetric lenses arise in synthesis.
Firstly, because each lens has a custom complement, we can no longer
specify the put functions though mere input/output examples.  Instead, we have
to specify the behavior through the observable behavior of edit sequences.
Furthermore, there are lenses that require arbitrarily long edit sequences to
differentiate between: consider the lens that keeps two formats equal for the
first $k$ non-identity edits, then keeps them completely unsynchronized for all
future edits.  This requires an edit sequence of $k$ edits to differentiate
between this lens and the identity lens.  Simple symmetric lenses avoid this
problem by not allowing such lenses: we can't keep track of the number of
non-identity edits performed unless there are explicit fields in the data
formats keeping track of them.  While we are reducing the expressivity of our
lenses with this restriction, we feel the benefits outweigh the costs.

@@

\paragraph*{Relationship with Asymmetric Lenses}
Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long2}:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}
\end{align*}
Despite being less expressive than symmetric lenses, simple symmetric lenses are
strictly more expressive than classical asymmetric lenses.

\begin{theorem}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem}
